Hop Probability Based Partitioning of Online Social Networks

### 2012-12-16 CMPT771/project_report/paper.tex
saving the last rivision

The paper is organized as follows. The related works is shown in section 2, the
objective function in exaplained in section 3, We describe our algorithm  in
section 4 and  experimental results are shown in section 5.  Finally, we
conclude the paper in section 6.

@@

    In this paper, we  compare our results with METIS\cite{20}. METIS focuses
on minimizing the cuts between the partitions as well as into roughly equal sized
partitions. The METIS algorithm is divided into three phases termed as
coarsening, partitioning and uncoarsening/refinement.  The main motivation
behind the algorithm is to generate an approximate graph of the original graph,
partitioning it and then retracting the coarsening phase while doing the
refinement between each iteration.  Our algorithm also follows a hierarchical
clustering approach where repeated merging of clusters is done for some
iteration and then the graph is randomly partitioned with imposed contraints of
having nearly equal partitions. Finally, the refinement algorithm is applied on
the clusters thus obtained. We differ from METIS, in terms of our partitioning
criterion, our novel way of clustering, and  the modified form Kerinighan-Lin 
algorithm which  refines the generated partitions.

\section{Hop probability Based Objective Function}

@@

\textbf{Phase 2 - Refining}\\
 Refine the partitions using kenighan-Lin algorithm or the modified one.\\
\hline
\end{tabular}
%\vskip -0.1in
\caption{Our Algorithm}
\label{alg}
\end{table*}

@@

The aim of this phase is to produce an approximate partitioning of the original
graph. Our approach is a heirarchical one.  We start with the original graph
and then we try to reduce the graph size by \emph{intelligently} clustering
nodes at each level. After clustering nodes, we form a new graph with clusters
from the previous level treated as nodes in the current level. After repeating
the steps for some iterations, we apply a traditional graph partitioning
algorithm to \emph{latest} graph. And then we do the refining on the produced
clusters to get the desired output. In the following sections, we will describe
our \emph{intelligent} clustering approach at each level and then we will
continue our discussion with the explanation of the refining algorithm.
\begin{itemize} % 1

@@

Graph  $G$=$(V,E)$, where $V$ = $\{v_1,v_2,v_3,\cdot\cdot\cdot,v_n\}$ ,$v_i$ represents a cluster of nodes. 
In the first level, $v_i$ consists of only single node and $E$ is the edge set which consists of $e_{ij}$ i.e the number of interactions between the $v_i$ and $v_j$ i.e for any $v_i$ and $v_j$,if there are $k$ edges between 
the two then $e_{ij}$ will be equal to $k$, for the first level $e_{ij}=number\ of\ interactions\ between\ v_i\ and\ v_j$.\footnote{number of interactions might the average number of times user $v_i$ and $v_j$ have interacted
with eachother}

@@

\subsection{Partitioning Phase}
After coarsening the graph, we can apply any traditional graph partitioning
approaches to partitiong it. In our case, we are randomly select nodes and put
them in partitions. While partitioning we also make sure that the size of does
not differ by large quantity i.e they should be of nearly equal size.
%\end{itemize} % 1

@@

    The Kernighan-Lin algorithm starts with initial bipartitions of the graph
and then iteratively optimizes it using a greedy approach. The Kernighan-Lin
algorithm is divided into two phases, in the first phase, we consider a pair of
nodes, with each nodes from different partitions. Now, a gain value $\delta$$Q$
is calculated for all possible pairs, which is the change in the number of
edges when the nodes in the pair are swapped. The pair with the maximum  gain
is marked for swapping so that they are never marked again in this phase. This
continues until all the pairs are marked. In the next phase, the algorithm
examine the sequence of the swapped pairs and the sequence which results into
the maximum gain is selected and the actual swapping takes place. This results
into a locally optimized graph and the two phases are repeated till there is no
further gain in the edges between the partitions.

    We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
To be precise, we change the definition of $\delta$$Q$ for us. In the
Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between
the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
in different partitions. Let $I_a$ be the internal cost for $a$, i.e the total
number of interactions between $a$ and its friends which lie in the same
partition and  $E_a$ be the external cost for the node $a$, i.e the total
number of interaction between $a$ and its friends which lie in the separate
partition. Similarily $E_b$ and $I_b$ are the external cost and internal cost
for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
- $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
  interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$. We
make changes in the internal, external and $C_{a,b}$ to fit our purpose of
decreasing the probability of the two hops from one node lying in the different
partitions. We define internal cost $I_{ap}$ as the probability of an edge of
node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
probability of an edge of node $a$ lying in a separate partition. Let $T_a$ be
$Total\ number\ of\ interactions\ done\ by\ user\ a$ and $T_b$ be $Total\
number\ of\ interactions\ done\ by\ user\ b$. And $I_{ab}$ be the number of
interactions between $a$ and $b$. Hence,\\
\begin{displaymath}I_{ap}=\frac{I_a}{T_a}\\\end{displaymath},
\\\begin{displaymath}E_{ap}=\frac{E_a}{T_a}\\\end{displaymath} and\\
\begin{displaymath}2C_{ap,bp}=\frac{I_{ab}}{T_a}+\frac{I_{ab}}{T_b}\\\end{displaymath}

    As the original Kernighan-Lin algorithm focuses on reducing the edges
between the partitions, our modified one reduces the probability of one hop of
the nodes to lie in the separate partition.  The rest of the kernighan-Lin
algorithm works in the original way. Experiments shows that it not only reduces
the probability of one hop to lie outside the partitions but also decreases the
edge cuts. The same approach could be applied to the linear heuristics from the
feduccia and matthews algorithm\cite{23} which can make it more efficient in
terms of time taken to refine the partitions.

@@

\label{fig:twopartcuts}
\end{table*}
\begin{table*}[htb]
\centering
\caption{Two Partitions - Evaluation Indexes}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
 &\multicolumn{2}{c|}{Our Algorithm}&\multicolumn{2}{c|}{METIS}\\
\hline
%\cline{2-5}
 DataSets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
\hline
Facebook& 0.225316 & 0.273173  & 0.21507 & 0.269943\\
\hline
SOC-Epinions & 0.135977 & 0.304285  & 0.162591 & 0.313931\\
\hline
SLASHDOT& 0.37827 & 0.496867  & 0.438342 & 0.762978\\
\hline
\end{tabular}
\label{fig:twopartindex}
\end{table*}

\section{Experiments and Results}
We have picked three popular and reasonaably sized datasets to run our experiments on
\begin{itemize}
\item \textbf{Facebook Social Network}: Facebook is a true social networking
site on which users interact by writing on the wallposts and sharing lots of
media content. The dataset was collected between December 29th, 2008 and
January 3rd, 2009, and between January 20th 2009 and January 22th 2009 from the
Facebook New Orleans network. The original dataset is time stamped, mentioning
when were the links made.  For our purposes we used it as an undirected graph
giving no consideration to time stamp attribute.  We also have list of all the
wall posts for Facebook New Orleans network. The dataset was anonymized and we
only have information about which user posted on whose wall at what time. No
posts content is provided. But this information was good enough to find the
interactions between the users which forms the weights of the edges between the
users.  So we build a graph in which nodes are represented by users and there's
and edge between two nodes if those two users are friends and weights are the
average number of interactions between the two users i.e average number of
times they posted on each other's wall + 1. 1 is added as there were cases when
two users were friends but they didn't interact with each other. So we decided
on having edge weight as 1 by default. As atleast they interacted once when
they accepted the friends request. The dataset was crawled by \cite{26}.

@@

\end{itemize}

We compare the hop probability based partitioning approach with METIS. We are
considering only bipartitions in our results. The input graph to the two
algorithms are same except the facebook graph. For our algorithm, the input
Facebook graph is with edge weights based on  interactions between users, while
input Facebook graph for METIS is with all the edge weights set to 1. The main
reason is that METIS doesn't take edge weights as input, so we can't provide it
with that. We also have a reasonable contraint that the partitions should be of
roughly equal size. We compare the two algorithms on edges cuts and hop
probability based evaluation criteria. As we can see the periliminary results
are quite close to METIS in terms of edge cuts as well the novel criteria. We
think that some more experimentation with the weight function, partitioning
algorithm(here we are doing random partitioning) and the refinement algorithm,
might result in better results. 
\section{Conclusion and Future Work}
In this paper, our focus was to partition an Online Social Network graph such
that the probability of query(one-hop/two-hop) generated from a user lying in
different partition is reduced. We proposed a novel criteria for evaluating the
partitions and presented a hierarchical algorithm optimizing that criteria. Our
preliminary results looks promising. We plan to further improve it by 
analyzing the graph formation at each level and also by experimenting with the
weight function as well as the partitioning function. Refinement algorithm can
be changed as well. As we don't know whether the modified Kernighan Lin
algorithm will surely converge. I plan to work on this a little more and maybe
submit it as a short paper in some conference/workshop.
\bibliographystyle{aaai} \bibliography{xbib}
\end{document}

>>>>

The paper is organized as follows. The related works is shown in section 2, the
objective function in explained in section 3, We describe our algorithm  in
section 4 and  experimental results are shown in section 5.  Finally, we
conclude the paper in section 6.

@@

    In this paper, we  compare our results with METIS\cite{20}. METIS focuses
on minimizing the cuts between the partitions as well as into roughly equal sized
partitions. The METIS algorithm is divided into three phases termed as
coarsening, partitioning and uncoarsening/refinement.  The main motivation
behind the algorithm is to generate an approximate graph of the original graph,
partitioning it and then retracting the coarsening phase while doing the
refinement between each iteration.  Our algorithm also follows a hierarchical
clustering approach where repeated merging of clusters is done for some
iteration and then the graph is randomly partitioned with imposed constraints of
having nearly equal partitions. Finally, the refinement algorithm is applied on
the clusters thus obtained. We differ from METIS, in terms of our partitioning
criterion, our novel way of clustering, and  the modified form Kernighan-Lin 
algorithm which  refines the generated partitions. In coarsening phase, our
algorithm tries to collapse the nodes into clusters so as to minimize the
evaluation criteria, while in METIS it doesn't take that into account. The refinement
phase also focuses on minimizing the evaluation criteria. 
\section{Hop probability Based Objective Function}

@@

\textbf{Phase 2 - Refining}\\
 Refine the partitions using Kernighan-Lin algorithm or the modified one.\\
\hline
\end{tabular}
%\vskip -0.1in
\caption{Our Algorithm}
\label{alg}
\end{table*}

@@

The aim of this phase is to produce an approximate partitioning of the original
graph. Our approach is a hierarchical one.  We start with the original graph
and then we try to reduce the graph size by \emph{intelligently} clustering
nodes at each level. After clustering nodes, we form a new graph with clusters
from the previous level treated as nodes in the current level. After repeating
the steps for some iterations, we apply a traditional graph partitioning
algorithm to \emph{latest} graph. And then we do the refining on the produced
clusters to get the desired output. In the following sections, we will describe
our \emph{intelligent} clustering approach at each level and then we will
continue our discussion with the explanation of the refining algorithm.
\begin{itemize} % 1

@@

Graph  $G$=$(V,E)$, where $V$ = $\{v_1,v_2,v_3,\cdot\cdot\cdot,v_n\}$ ,$v_i$ represents a cluster of nodes. 
In the first level, $v_i$ consists of only single node and $E$ is the edge set which consists of $e_{ij}$ i.e the number of interactions between the $v_i$ and $v_j$ i.e for any $v_i$ and $v_j$,if there are $k$ edges between 
the two then $e_{ij}$ will be equal to $k$, for the first level $e_{ij}=number\ of\ interactions\ between\ v_i\ and\ v_j$.\footnote{number of interactions might the average number of times user $v_i$ and $v_j$ have interacted
with each other}

@@

\subsection{Partitioning Phase}
After coarsening the graph, we can apply any traditional graph partitioning
approaches to partition it. In our case, we are randomly select nodes and put
them in partitions. While partitioning we also make sure that the size of does
not differ by large quantity i.e they should be of nearly equal size.
%\end{itemize} % 1

@@

    The Kernighan-Lin algorithm starts with initial Bi-partitions of the graph
and then iteratively optimizes it using a greedy approach. The Kernighan-Lin
algorithm is divided into two phases, in the first phase, we consider a pair of
nodes, with each nodes from different partitions. Now, a gain value $\delta$$Q$
is calculated for all possible pairs, which is the change in the number of
edges when the nodes in the pair are swapped. The pair with the maximum  gain
is marked for swapping so that they are never marked again in this phase. This
continues until all the pairs are marked. In the next phase, the algorithm
examine the sequence of the swapped pairs and the sequence which results into
the maximum gain is selected and the actual swapping takes place. This results
into a locally optimized graph and the two phases are repeated till there is no
further gain in the edges between the partitions.

    We modify one of the phases of  Kernighan-Lin algorithm for our criterion.
To be precise, we change the definition of $\delta$$Q$ for us. In the
Kernighan-Lin algorithm, $\delta$$Q$ is the change in the edges cuts between
the partitions when the nodes are swapped. Let $a$, $b$ be two nodes that lie
in different partitions. Let $I_a$ be the internal cost for $a$, i.e the total
number of interactions between $a$ and its friends which lie in the same
partition and  $E_a$ be the external cost for the node $a$, i.e the total
number of interaction between $a$ and its friends which lie in the separate
partition. $C_{a,b}$ is the average number of interactions
between user $a$ and $b$. Similarily $E_b$ and $I_b$ are the external cost and internal cost
for the node $b$. Now let $D_a$ = $E_a$ - $I_a$ and $D_b$ = $E_b$
- $I_b$ , i.e the difference of external and internal costs. If $a$ and $b$ are
  interchanged then the gain i.e $\delta$$Q$ will be $D_a + D_b -2C_{a,b}$. We
make changes in the internal, external and $C_{a,b}$ to fit our purpose of
decreasing the probability of the two hops from one node lying in the different
partitions. We define internal cost $I_{ap}$ as the probability of an edge of
node $E_{ap}$ lying in the same partition as of $a$ ,external cost as
probability of an edge of node $a$ lying in a separate partition. Let $T_a$ be
$Total\ number\ of\ interactions\ done\ by\ user\ a$ and $T_b$ be $Total\
number\ of\ interactions\ done\ by\ user\ b$. And $I_{ab}$ be the number of
interactions between $a$ and $b$. Hence,\\
\begin{displaymath}I_{ap}=\frac{I_a}{T_a},\\\end{displaymath}
\\\begin{displaymath}E_{ap}=\frac{E_a}{T_a}\\\end{displaymath} and\\
\begin{displaymath}2C_{ap,bp}=\frac{I_{ab}}{T_a}+\frac{I_{ab}}{T_b}\\\end{displaymath}

    As the original Kernighan-Lin algorithm focuses on reducing the edges
between the partitions, our modified one reduces the probability of one hop of
the nodes to lie in the separate partition.  The rest of the Kernighan-Lin
algorithm works in the original way. Experiments shows that it not only reduces
the probability of one hop to lie outside the partitions but also decreases the
edge cuts. The same approach could be applied to the linear heuristics from the
feduccia and Mathews algorithm\cite{23} which can make it more efficient in
terms of time taken to refine the partitions.

@@

\label{fig:twopartcuts}
\end{table*}
\begin{table*}[htb]
\centering
\caption{Two Partitions - Evaluation Indexes}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
 &\multicolumn{2}{c|}{Our Algorithm}&\multicolumn{2}{c|}{METIS}\\
\hline
%\cline{2-5}
 Datasets&Hop Probability score &Edge Cuts&Hop Probability Score&Edge Cuts\\
\hline
Facebook& 0.042590 & 87642  & 0.031047 & 68315\\
\hline
SOC-Epinions & 0.192315 & 62544  & 0.163753 & 55104\\
\hline
SLASHDOT& 0.398257 & 113356  & 0.438645 & 81465 \\
\hline
\end{tabular}
\label{fig:twopartindex}
\end{table*}

\section{Experiments and Results}
We have picked three popular and reasonably sized datasets to run our experiments on
\begin{itemize}
\item \textbf{Facebook Social Network}: Facebook is a true social networking
site on which users interact by writing on the wall-posts and sharing lots of
media content. The dataset was collected between December 29th, 2008 and
January 3rd, 2009, and between January 20th 2009 and January 22th 2009 from the
Facebook New Orleans network. The original dataset is time stamped, mentioning
when were the links made.  For our purposes we used it as an undirected graph
giving no consideration to time stamp attribute.  We also have list of all the
wall posts for Facebook New Orleans network. The dataset was anonymized and we
only have information about which user posted on whose wall at what time. No
posts content is provided. But this information was good enough to find the
interactions between the users which forms the weights of the edges between the
users.  So we build a graph in which nodes are represented by users and there's
and edge between two nodes if those two users are friends and weights are the
average number of interactions between the two users i.e average number of
times they posted on each other's wall + 1. 1 is added as there were cases when
two users were friends but they didn't interact with each other. So we decided
on having edge weight as 1 by default. As at-least they interacted once when
they accepted the friends request. The dataset was crawled by \cite{26}. We have chosen
this dataset as it is one of the few Facebook datasets provided on the web, which also has
the wall-posts information during that period. That gives us a close approximate of
the users interaction amongst each other.

@@

\end{itemize}
We compare the hop probability based partitioning approach with METIS. We are
considering only Bi-partitions in our results. The input graph to the two
algorithms are same except the facebook graph. For our algorithm, the input
Facebook graph is with edge weights based on  interactions between users, while
input Facebook graph for METIS is with all the edge weights set to 1.
We also have a reasonable constraint that the partitions should be of
roughly equal size. We compare the two algorithms on edges cuts and hop
probability based evaluation criteria.

As we can see the preliminary results
are quite close to METIS in terms of edge cuts as well the novel criteria. We
think that some more experimentation with the weight function, partitioning
algorithm(here we are doing random partitioning) and the refinement algorithm,
might result in better results. Given the time span for the course project,
we thought of running our algorithm on relatively small graphs. We wanted to check the performance
in terms of edge cuts and the probability based evaluation criteron. We feel that
since the first iteration of the algorithm gave us quite close results, we beleive further improvement
over them will beat the more refined $METIS$ algorithm which took few years to develop.

\section{Conclusion and Future Work}
In this paper, our focus was to partition an Online Social Network graph such
that the probability of query(one-hop/two-hop) generated from a user lying in
different partition is reduced. We proposed a novel criteria for evaluating the
partitions and presented a hierarchical algorithm optimizing that criteria. Our
preliminary results looks promising. We plan to further improve it by 
analyzing the graph formation at each level and also by experimenting with the
weight function as well as the partitioning function. Refinement algorithm can
be changed as well. As we don't know whether the modified Kernighan Lin
algorithm will surely converge. We also want to compare the algorithm 
with variety of graph partitioning algorithm other than METIS and also design 
an algorithm for generating $k$ partitions. I plan to work on this a little more and 
 eventually submit it as a workshop paper in ICWSM'13.
\bibliographystyle{aaai} \bibliography{xbib}
\end{document}
